Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-05 trunk/paper_enahpc14/Sec3:Environment-setup.tex
updates working copy

\begin{itemize}
\item \monch is  a 10-rack NEC-provided clus\-ter at  CSCS composed of
  312 compute nodes. 
  % and  utilized  by  members  of  the  Swiss  Platform  for  Advanced
  % Scientific Computing (PASC) project~\cite{PASC}.
  Each node is  comprised of two Intel Ivy  Bridge EP E5-2660 ten-core
  processors  operating  at 2.2\,GHz,  equipped  with  32\,GB of  DDR3
  1600\,MHz RAM  and connected via InfiniBand Mellanox  SX6036 and FDR
  switches. For our  experiments, a full rack constituted  of 52 nodes
  was used. 
  % \monch is  slated to stay  in service without hardware  upgrade for
  % the  duration  of  the  Exa2Green project,  allowing  an  identical
  % configuration  to  be  recreated  for  future  assessments  of  the
  % baseline.

\item \pilat is  a dual-socket eight-core Intel Sandy  Bridge EP based
  cluster at CSCS composed of 42 compute nodes.  
  % and used as Piz Daint  pre-post processing cluster The 2 login and
  % 42  compute  nodes  are   consolidated  into  11  twin-pair  Intel
  % E5-Series  DALCO  r2264i4t   2U  scalable  compute  modules,  each
  % containing 4 compute nodes
  Each node  is comprises two Intel Xeon  E5-2670 processors operating
  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected
  via  InfiniBand   Mellanox  SX6036   and  FDR  switches.    For  our
  experiments, a full  rack constituted of 42 compute  nodes was used.
  \pilat   is   based   on   Intel's   second-generation   processors,
  conventional in  HPC systems  and known to  be more  power consuming
  than its Ivy Bridge successor.

@@

For  experiments  at UJI  on  \tinto, we  employed  a  version of  the
integrated  framework presented  in~\cite{energy13}  that employs  the
\pmlib  library in  combination  with Extrae  and  Paraver, which  are
profiling/tracing and visualization tools, respectively.

To employ our  framework, \cosmoart is first compiled  and linked with
the  Extrae instrumentation  library, which  automatically instruments
the Fortran~90  code of the model,  e.g., to record  MPI calls.  Next,
\cosmoart  is run  on the  nodes, thus  dissipating certain  amount of
power. The  nodes are connected  to APC 8653 Power  Distribution Units
(PDUs) that account  for the dissipated power with  a sampling rate of
1\,Hz. The power data captured by  these PDUs is collected by a \pmlib
tracing server running  on a separate machine. The  client, running on
the target nodes, sends start/stop  primitives in order to collect the
power  data during  the \cosmoart  run. Once  the run  is  finished, a
power-performance  trace file  containing the  power data  is received
from  the  tracing  server  and the  instrumentation  post-process  is
generated.  A  merged trace  file  is  finally  loaded in  Paraver  to
visualize  and correlate  the  performance trace  and  with the  power
profile  generated by \cosmoart.  This procedure  allows us  to easily
detect power bottlenecks, basically constituted by inefficient waiting
methods.

@@

A snapshot of the code, which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced  and run on 1040 cores  using 20~MPI tasks per
node on  \monch, on 1344 cores  using 16~MPI tasks per  node on \pilat
and  on  192  cores  using  12~MPI  tasks per  node  on  \tinto.   The
calculated region  was mapped to the participating  processors using a
2D-partitioning  strategy.  The  distribution  along the  $x$ and  $y$
coordinates  was   defined  accordingly  depending   on  the  selected
platform.  Besides,  as  this  version   does  not  make  use  of  the
traditional  GRIB  library, we  specified  $nprocio=0$  for GRIB  I/O.
Hyperthreading  was not  considered  in this  study  as it  previously
revealed that it always led to higher energy-to-solution.

>>>>

\begin{itemize}
\item \monch is  a 10-rack NEC-provided clus\-ter at  CSCS composed of
  312 compute nodes. 
  % and  utilised  by  members  of  the  Swiss  Platform  for  Advanced
  % Scientific Computing (PASC) project~\cite{PASC}.
  Each node is  comprised of two Intel Ivy  Bridge EP E5-2660 ten-core
  processors  operating  at 2.2\,GHz,  equipped  with  32\,GB of  DDR3
  1600\,MHz RAM  and connected via InfiniBand Mellanox  SX6036 and FDR
  switches. For our  experiments, a full rack constituted  of 52 nodes
  was used. 
  % \monch is  slated to stay  in service without hardware  upgrade for
  % the  duration  of  the  Exa2Green project,  allowing  an  identical
  % configuration  to  be  recreated  for  future  assessments  of  the
  % baseline.

\item \pilat is  a dual-socket eight-core Intel Sandy  Bridge EP based
  cluster at CSCS composed of 42 compute nodes.  
  % and used as Piz Daint  pre-post processing cluster The 2 login and
  % 42  compute  nodes  are   consolidated  into  11  twin-pair  Intel
  % E5-Series  DALCO  r2264i4t   2U  scalable  compute  modules,  each
  % containing 4 compute nodes based on two ...
  Each node  is comprises two Intel Xeon  E5-2670 processors operating
  at 2.6\,GHz equipped with 64\,GB of DDR3 1600\,MHz RAM and connected
  via  InfiniBand   Mellanox  SX6036   and  FDR  switches.    For  our
  experiments, a full  rack constituted of 42 compute  nodes was used.
  \pilat   is   based   on   Intel's   second-generation   processors,
  conventional in  HPC systems  and known to  be more  power consuming
  than its Ivy Bridge successor.

@@

For  experiments  at UJI  on  \tinto, we  employed  a  version of  the
integrated  framework presented  in~\cite{energy13}  that employs  the
\pmlib  library in  combination  with Extrae  and  Paraver, which  are
profiling/tracing and visualisation tools, respectively.

To employ our  framework, \cosmoart is first compiled  and linked with
the  Extrae instrumentation  library, which  automatically instruments
the Fortran~90  code of the model,  e.g., to record  MPI calls.  Next,
\cosmoart  is run  on the  nodes, thus  dissipating certain  amount of
power. The  nodes are connected  to APC 8653 Power  Distribution Units
(PDUs) that account  for the dissipated power with  a sampling rate of
1\,Hz. The power data captured by  these PDUs is collected by a \pmlib
tracing server running  on a separate machine. The  client, running on
the target nodes, sends start/stop  primitives in order to collect the
power  data during  the \cosmoart  run. Once  the run  is  finished, a
power-performance  trace file  containing the  power data  is received
from  the  tracing  server  and the  instrumentation  post-process  is
generated.  A  merged trace  file  is  finally  loaded in  Paraver  to
visualise  and correlate  the  performance trace  and  with the  power
profile  generated by \cosmoart.  This procedure  allows us  to easily
detect power bottlenecks, basically constituted by inefficient waiting
methods.

@@

A snapshot of the code, which includes, at least conceptually, all the
information needed  to reproduce the  energy-to-solution benchmarks of
\cosmoart, was produced  and run on 1040 cores  using 20~MPI tasks per
node on  \monch, on 1344 cores  using 16~MPI tasks per  node on \pilat
and  on  192  cores  using  12~MPI  tasks per  node  on  \tinto.   The
calculated region  was mapped to the participating  processors using a
2D-partitioning  strategy.  The  distribution  along the  $x$ and  $y$
coordinates was  defined by  setting: $nprocx=40$ and  $nprocy=26$ for
\monch,  $nprocx=28$ and  $nprocy=24$ for  \pilat and  $nprocx=??$ and
$nprocy=??$  for \tinto as  $nprocx$ is  usually kept  slightly bigger
than  $nprocy$. Besides,  as this  version does  not make  use  of the
traditional  GRIB  library, we  specified  $nprocio=0$  for GRIB  I/O.
Hyperthreading  was not  considered  in this  study  as it  previously
revealed that it always led to higher energy-to-solution.
